<!-- -*-markdown-*- -->
# stevedore

[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![Travis-CI Build Status](https://travis-ci.org/richfitz/stevedore.svg?branch=master)](https://travis-ci.org/richfitz/stevedore)
[![codecov.io](https://codecov.io/github/richfitz/stevedore/coverage.svg?branch=master)](https://codecov.io/github/richfitz/stevedore?branch=master)

A docker client for R

**WARNING**: package is in early development. If you run the test suite it may do all sorts of things to your containers/images/volumes/networks.  Do not run it unless you'd be happy running `docker system prune -f`

```{r, prep, include = FALSE}
knitr::opts_chunk$set(error = FALSE)
has_internet <- function() {
  !is.null(suppressWarnings(utils::nsl("www.google.com")))
}
if (has_internet()) {
  d <- stevedore::docker_client()
  try(d$images$remove("alpine:3.1"), silent = TRUE)
}
```

## Usage

The main function in the package is `docker_client`; this will construct an object with which we can talk with the docker server.

```{r, create}
docker <- stevedore::docker_client()
docker
```

With this you can run containers:

```{r, run}
docker$containers$run("alpine:3.1", c("echo", "hello world"))
```

Or run containers in the background

```{r, background}
docker$containers$run("bfirsh/reticulate-splines", detach = TRUE)
```

You can manage containers

```{r, list}
docker$containers$list()
id <- docker$containers$list(limit = 1L)$id
container <- docker$containers$get(id)
container
```

And control containers

```{r, control}
container$inspect()$config$image
container$logs()
container$stop(t = 0)
container$remove()
```

And manage images

```{r, images}
head(docker$images$list())
```

## Approach

Docker publishes a [machine-readable API specification](https://docs.docker.com/engine/api/v1.29).  Rather than manually write wrappers that fit the output docker gives, `stevedore` _generates_ an interface directly from the spefification.  Currently `stevedore` supports docker API versions `r stevedore:::MIN_DOCKER_API_VERSION` to `r stevedore:::MIN_DOCKER_API_VERSION` (defaulting to `r stevedore:::DEFAULT_DOCKER_API_VERSION`).

This approach means that the output will be type-stable - there is no inference on what to return based on what the server chooses to return.  With a given API version, the same fields will always be returned.  Some of this information is very rich, for example, for the backgrounded container above:

```{r, inspect}
container$inspect(reload = FALSE)
```

## Roadmap

There is still a lot of work to do here:

* windows support needs work - the current approach works only for a unix socket
* unix non-socket (tcp) access, and TLS
* endpoints that require http hijacking are not fully supported (attach, logs) but the foundations are there to support this - stdin is likely to be a major hassle though
* endpoints that require tar input and output (equivalents of `docker cp` especially) need major work
* lots of work on parameter wrangling for the more complicated endpoints (basically things that take anything more complicated than a string array are prone to failure because I've not tested them yet)
* swarm features (`nodes`, `plugins`, `secrets`, `services` and `swarm`) are not implemented - not because they'd be any harder but just because I've never used them
* authentication of registries for pull and push (not hard, just not done)
* some sort of documentation - hard given that the interface is directly pulled from the spec so we need to patch up references to json

## Licence

MIT © [Rich FitzJohn](https://github.com/richfitz).

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
