---
title: "Cookbook"
author: "Rich FitzJohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cookbook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

``` {r include = FALSE}
knitr::opts_chunk$set(error = FALSE)
lang_output <- function(x, lang) {
  cat(c(sprintf("```%s", lang), x, "```"), sep="\n")
}
c_output <- function(x) lang_output(x, "cc")
r_output <- function(x) lang_output(x, "r")
plain_output <- function(x) lang_output(x, "plain")
```

This vignette contains short snippets of `stevedore` calls to
perform simple tasks, alongside a docker command line equivalent.

Conventions:

* We assume that a `stevedore` docker client object has been
  created with `docker <- stevedore::docker_client()` (to avoid
  repetition in examples)
* All docker command line examples start with `$` to represent the
  unix/bash prompt

``` {r }
```{r}
docker <- stevedore::docker_client()
```
```

## Start a container with ports published to the host

``` {r }
```{r, include = FALSE}
docker$image$pull("gplates/gws")
```
```

With docker cli:

```
$ docker run --rm -p 8888:80 gplates/gws
```

the `-p 8888:80` publishes port `80` in the container to port
`8888` on the host.

In `stevedore`, this can be done with:
``` {r }
gws <- docker$container$run("gplates/gws", ports = "8888:80",
                            detach = TRUE, rm = TRUE)
gws$ports()
```

``` {r include = FALSE}
gws$kill()
```

To use a random port, omit the host portion (here `8888:`) as

```
$ docker run --rm -p 80 gplates/gws
```

or, with `stevedore`
``` {r }
gws <- docker$container$run("gplates/gws", ports = 80,
                            detach = TRUE, rm = TRUE)
```

To determine the port you can use the `ports()` method
``` {r }
gws$ports()
```

``` {r include = FALSE}
gws$kill()
```

If the container declares its exposed ports using an `EXPOSE
<port>` directive in the `Dockerfile` then you can automatically
publish all declared ports to random host ports with `-P` as

```
$ docker run --rm -P gplates/gws
```

or with `stevedore` by using `ports = TRUE`

``` {r }
gws <- docker$container$run("gplates/gws", ports = TRUE, detach = TRUE)
gws$ports()
```

``` {r include = FALSE}
gws$kill()
```

<small>As contributed by [Brian O'Meara](https://github.com/bomeara) in
[#44](https://github.com/richfitz/stevedore/issues/44)</small>
