---
title: "Using docker with stevedore"
author: "Rich FitzJohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using docker with stevedore}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`stevedore` is an R package for interacting with docker from R.
With `stevedore` you can

* create, run, manage, remove and destroy containers
* build, pull, manage and remove images
* create, destroy and interact with docker networks
* create, destroy and interact with docker volumes

Almost everything that can be done with the docker command line
client (the main exception being that is not possible to send input
to a container as if you were on a terminal).

`stevedore` directly interacts with the docker server over its HTTP
API.  This gives a more direct access to docker than going via the
command line.

This vignette quickly walks through the core features of the
package.  Because `stevedore` wraps the docker API directly there
are many more arguments that can be used than are covered here.
But this covers the core use.

``` {r include = FALSE}
knitr::opts_chunk$set(error = FALSE)
lang_output <- function(x, lang) {
  cat(c(sprintf("```%s", lang), x, "```"), sep="\n")
}
c_output <- function(x) lang_output(x, "cc")
r_output <- function(x) lang_output(x, "r")
plain_output <- function(x) lang_output(x, "plain")
build_iterate <- function() {
  p <- system.file("images/iterate", package = "stevedore", mustWork = TRUE)
  stopifnot(file.copy(p, ".", recursive = TRUE))
  stevedore::docker_client()$image$build("iterate", tag = "richfitz/iterate")
}
```
``` {r include = FALSE}
build_iterate()
nginx_ready <- function(port, attempts = 10) {

  f <- function() {
    url <- sprintf("http://localhost:%s", port)
    curl::parse_headers(curl::curl_fetch_memory(url)$headers)
    TRUE
  }
  for (i in seq_len(attempts)) {
    if (tryCatch(f(), error = function(e) FALSE)) {
      return()
    }
    Sys.sleep(1)
  }
  stop("nginx not available in time")
}
```



The main function in the package is `docker_client`; this will
construct an object with which we can talk with the docker server.
By default this will look at a number of environment variables and
try to connect to the correct daemon.  See `?docker_client` for
information on controlling creating the connection.
``` {r }
docker <- stevedore::docker_client()
```

The client object looks a lot like an
[`R6`](https://github.com/wch/R6) object (though it is implemented
differently because the interface here is automatically generated
in ways that don't play nicely with R6).  But if you are at all
familiar with R6 objects it should seem quite familiar.
``` {r }
docker
```

Each function call (e.g., `ping`) is callable by accessing with
`$`, such as
``` {r }
docker$ping()
```

In addition there are "collection" objects (e.g., `container`)
that are accessed using `$`, like a directory structure
``` {r }
docker$container
```

The interface is designed similarly to the command line docker
client (and to the Python docker client), where container commands
are within the `container` collection and image commands are within
the `image` collection and so on (the main difference with the
command line client is that the container commands are not at the
top level, so it is `docker$container$run(...)` not
`docker$run(...)`).

To run a container, the `docker$container$run` command follows the
semantics of the command line client and will

* pull a image if it does not exist
* create the container
* run the container
* fetch the logs
``` {r }
res <- docker$container$run("hello-world")
```

This returns a list with two elements:
``` {r }
names(res)
```

The "logs" element is the logs themselves:
``` {r }
res$logs
```

This is a `docker_stream` object and codes the output stream using
the `stream` attribute (it can otherwise be treated as a character
vector).

The "container" element is an object that can be used to interact
with a container
``` {r }
res$container
```

For example the function `path_stat` gets some information about
paths on the container:
``` {r }
res$container$path_stat("hello")
```

The image can also be returned
``` {r }
img <- res$container$image()
img
```

which is another object with methods that can be invoked to find
out about the image, e.g.:
``` {r }
img$history()
```

## Containers

The `$container` object includes methods for interacting with
containers:
``` {r }
docker$container
```

`$create` creates a new container (similar to `docker container
create` on the command line) but does not start it.
``` {r }
x <- docker$container$create("hello-world", name = "hello-stevedore")
x
```

`$get` creates a `docker_container` object from an container id or
name:
``` {r }
y <- docker$container$get("hello-stevedore")
x$id()
y$id()
```

`$list()` lists containers (like `docker list` on the command line)
- by default showing only running containers
``` {r }
docker$container$list()
docker$container$list(all = TRUE, limit = 2)
```

`$remove()` removes a container by name or id:
``` {r }
docker$container$remove("hello-stevedore")
```

`$prune()` removes non-running containers (i.e., containers that
have exited or containers that have been created but not yet
started)
``` {r }
docker$container$prune()
```

### Working with containers

After creating a container object, there are many more methods to
use - all apply to the individual container
``` {r }
x <- docker$container$create("richfitz/iterate", c("1000", "1"))
```

Most are analogues of similarly named `docker` command line
functions.

First, there are some basic query methods - `$id()`, `$name()` and
`$labels()`
``` {r }
x$id()
x$name()
x$labels()
```

More detailed information (**much** more detailed) can be retrieved
with the `$inspect()` method
``` {r }
x$inspect()
```

The image used by a container can be retrieved with the `$image()`
method
``` {r }
x$image()
```

(see below for working with images).

The status of the container (`created`, `running`, `exited`,
`paused`, etc) can be read with `$status()`
``` {r }
x$status()
```

The container created by by `$create` is not running - the
`$start()` method will start it:
``` {r }
x$start()
x$status()
```

Once the container is running we can query to see what processes
are running in it with `$top` (standing for Table Of Processes)
``` {r }
x$top()
```

We can also get the logs:
``` {r }
x$logs()
```

This returns a special object type `docker_stream` which allows
control over formatting with `format()` - the `style` argument
controls how stderr and stdout are printed.  There is a `stream`
attribute that can be used to separate out lines too.  If a tty was
allocated with `tty = TRUE` the output will be a plain character
vector
``` {r }
format(x$logs(), style = "plain")
```

It can generally be treated as a character vector:
``` {r }
x$logs()[1:2]
```

The `$logs()` method can be used to do a blocking wait on a
container.  Pass `follow = TRUE` to follow the logs.  You will want
to provide a `stream` argument too, which is where to stream the
log to.  This can be `stdout()` or `stderr()`, a file or an R
connection.
``` {r }
y <- docker$container$create("richfitz/iterate", c("10", "0.1"))
y$start()
y$logs(stream = stdout(), follow = TRUE)
```

If running this interactively, the logs will print one line at a
time - once control returns to R the container has exited.  You can
escape this streaming using whatever method you use to interrupt an
R calculation (depends on which GUI/IDE you are using) but the
container will continue regardless - we are just _observing_ a
running container.
``` {r }
y$status()
```

The other way of blocking until a container has finished is with
`$wait()` which blocks until the container exits, then returns the
exit code.
``` {r }
y$start()
y$wait()
```

Calling `$wait()` on an exited container is fine, and will just
return immediately:
``` {r }
y$wait()
```

Containers can be paused with `$pause()`
``` {r }
x$pause()
x$status()
```

Once paused, they can be restarted with `$unpause()`
``` {r }
x$unpause()
x$status()
```

Additionally, containers can be _restarted_ with `$restart()`
``` {r }
x$restart(t = 0)
x$logs(tail = 5)
```

Containers can be stopped with `$stop()` and removed with
`$remove()` (calling `$remove(force = TRUE)` will kill the
container before removing.
``` {r }
x$stop(t = 0)
x$remove()
```

Once a container has been removed most methods will not work properly:
``` {r error = TRUE}
x$status()
```

``` {r include = FALSE}
y$remove()
```

Information about ports (for containers that expose them) can be
retrieved with `$ports()`.  The `nginx` image creates a web
server/proxy that exposes port 80 from the container.  We can map
that to a random port by asking docker to expose port 80 but not
saying what to map it to:
``` {r }
nginx <- docker$container$run("nginx", ports = 80,
                              detach = TRUE, rm = TRUE,
                              name = "stevedore-nginx")
nginx$ports()
```

(alternatively, use `ports = TRUE` to act like `docker run`'s `-P`
and "publish all ports to random ports).

``` {r include=FALSE}
nginx_ready(nginx$ports()$host_port)
```

This shows that the port exposed by the the container (80) is
mapped to the port `r as.integer(nginx$ports()$host_port)` on the
host.  We can use this to communicate with the server:
``` {r }
url <- sprintf("http://localhost:%s", nginx$ports()$host_port)
curl::parse_headers(curl::curl_fetch_memory(url)$headers)
nginx$stop(t = 0)
```


## Run commands in running containers with `exec`

With the command-line tool, `docker exec <container name>
<command>` lets you run an arbitrary command within a running
container.  `stevedore` does this with the `exec` method of a
container object.

Reasons for doing this include debugging (using arbitrary commands
to inspect/interact with a container while it does its primary
task) but it can also be used in deployment (e.g., sending a "go"
signal after copying files into the container).

To demonstrate, we need a long running container:
``` {r }
x <- docker$container$run("richfitz/iterate", c("1000", "10"),
                          detach = TRUE, rm = TRUE)
x$status()
```

With the container running we can run additional commands:
``` {r }
res <- x$exec("ls")
```

This streams the output of the command by default (to the
connection indicated by the `stream`) argument.  Output is also
returned as part of the object:
``` {r }
res
```

## Copy files into and out of containers

Just like `docker cp`, `stevedore` lets you copy files into and out
of containers.  The logic mimics the logic in the docker command
line client as closely as possible.

To copy a file into our container `x`, use `$cp_in`
``` {r }
path <- tempfile()
writeLines("hello", path)
x$cp_in(path, "/hello")
```

And the new file is on the container
``` {r }
x$exec(c("cat", "/hello"), stream = FALSE)$output
```

The input can be a single file or a single directory.

To copy out, use `$cp_out`
``` {r }
dest <- tempfile()
x$cp_out("/usr/local/bin/iterate", dest)
```

Here is the iterate script, from the container:
``` {r }
writeLines(readLines(dest, n = 10))
```

There is also a convenience method at the root of the docker object
that behaves more like `docker cp` and requires that one of the source or destination arguments is given in `<container>:<path>` format:
``` {r }
src <- paste0(x$name(), ":/usr/local/bin/iterate")
src
```

as
``` {r }
dest2 <- tempfile()
docker$cp(src, dest2)
```

which achieves the same thing as the `$cp_out` command above.
``` {r }
unname(tools::md5sum(c(dest, dest2)))
```

(don't forget to remove your detached containers later!)
``` {r }
x$kill()
```

## Images

### Pulling

``` {r include = FALSE, error = TRUE}
try(docker$image$remove("bash:latest"), silent = TRUE)
```

Images can be directly pulled with `docker$image$pull` providing
an image name (as either `<repo>` or `<repo>:<tag>`.  If the image
exists already this will be quick, and if the network connection is
down then this will fail.
``` {r }
docker$image$pull("bash:latest")
```

### The object returned by pull is an image object - that can be
### created by using `$get`
``` {r }
img <- docker$image$get("bash:latest")
```

### Building

The other common way of getting images is to build them (the
equivalent of `docker build`).  So if we have a path (here,
`iterate`) containing a Dockerfile:
``` {r }
dir("iterate")
```

The Dockerfile itself contains:
``` {r echo = FALSE, results = "asis"}
plain_output(readLines("iterate/Dockerfile"))
```

and the `iterate` file is an executable shell script containing:
``` {r echo = FALSE, results = "asis"}
lang_output(readLines("iterate/iterate"), "shell")
```

We can build this image using:
``` {r }
img <- docker$image$build("iterate", tag = "richfitz/iterate", nocache = TRUE)
img
```

The newly created image is returned as an image object and can be
used via `$container$run()`
``` {r }
invisible(
  docker$container$run(img, c("10", "0.1"), rm = TRUE, stream = stdout()))
```

``` {r include = FALSE}
unlink("iterate", recursive = TRUE)
```

### Importing

There is a third way of creating an image, which is to import it
from a tar archive.  This is not yet documented (**TODO**) but can
be done via `$image$import()`

### Working with image objects

Each image object has a number of methods.
``` {r }
img <- docker$image$get("richfitz/iterate")
img
```

The `$id()`, `$short_id()`, `$labels()`, `$name()` and `$tags()`
query basic information about an image
``` {r }
img$id()
img$short_id()
img$labels()
img$name()
img$tags()
```

(short_id` is always 10 characters long and does not include a
leading `sha256:`).

The `inspect()` method returns detailed information about the image
``` {r }
img$inspect()
```

the exact format varies between docker API versions but should be
the same for all images within an API version.

The `history()` method returns a data.frame of information about
the history of an image (i.e., the layers that it is constructed
from)
``` {r }
img$history()
```

(printing these objects is a real challenge!).

The `export()` method exports an image as a tar object.  There is
some work still required to make this work nicely (currently it
returns a [potentially long] raw vector).

There are several methods that operate to modify or destroy the image:

The `$tag()` method will tag an image, for example
``` {r }
img$tag("richfitz/iterate", "0.0.1")
img$reload()
img$tags()
```

While the `$untag()` method will remove a tag
``` {r }
img$untag("richfitz/iterate:0.0.1")
```

The `$remove()` method will remove an image - this returns a
data.frame indicating what actually happened (images are only
actually deleted if there are no other tags pointing at an image
_and_ if `noprune` is not `TRUE`.
``` {r }
img$remove()
```

``` {r include = FALSE}
build_iterate()
```

## Volumes

Docker volumes provide a useful abstraction for interacting with
(possibly persistent) file volumes across containers.  To create a
volume using `stevedore` (equivalent to `docker volume create`) use
`$volume$create()`:
``` {r }
vol <- docker$volume$create("myvolume")
```

Volumes can be listed:
``` {r }
docker$volume$list()
```

### Working with volume objects

There's very little that can be done with volume objects:
``` {r }
vol
```

We can get the name:
``` {r }
vol$name()
```

Inspect the metadata
``` {r }
vol$inspect()
```

Generate mount definitions:
``` {r }
vol$map("/comtainer/path")
```

and can remove the volume
``` {r }
vol$remove()
```

everything else comes from using volumes with containers.

Containers are mounted the same way as from the docker command line
- with a string in the form `<host>:<container>`.  Here we can use
the volume name on the host side, so saying `<myvolume>:/myvolume`
mounts our volume at `/myvolume` within the container.  This can be
done easily with the `$map()` method.
``` {r }
vol <- docker$volume$create("myvolume")
docker$container$run(
  "alpine:latest",
  c("sh", "-c", "echo hello world > /myvolume/output"),
  volumes = vol$map("/myvolume"),
  rm = TRUE)
```

(We use `sh -c` here so that the redirect operates within the
container - the third argument is evaluated by the shell within the
container and redirects the value that is echoed to a file.)

We can see the result of this by using a second container to read
the file:
``` {r }
docker$container$run(
  "alpine:latest",
  c("cat", "/myvolume/output"),
  volumes = vol$map("/myvolume"),
  rm = TRUE)$logs
```

``` {r include = FALSE}
vol$remove()
```

## Networks

Docker "networks" make it easy to get containers communicating with
each other without exposing ports to the host.  To achive this, one
creates a docker network, then create containers attached to that
network (containers can also be attached to networks after
creation).
``` {r }
nw <- docker$network$create("mynetwork")
```

Networks can be listed:
``` {r }
docker$network$list()
```

The networks `bridge`, `host` and `none` always exist - they are
special to docker.

### Working with network objects

Like volume objects, network objects do very little themselves:
``` {r }
nw
```

The `$name()` and `$id()` methods get the name and id of the network
``` {r }
nw$name()
nw$id()
```

`$inspect()` gets detailed metadata
``` {r }
nw$inspect()
```

`$containers()` lists containers attached to the network (currently
an empty list - this network has no attached containers)
``` {r }
nw$containers()
```

`$remove()` removes the network
``` {r }
nw$remove()
```

Generally you'll want to put containers onto a network.

The setup here is to create a network, and then use the `network`
argument to `$container$run()` to attach a container to that
network.  Once established, containers on the same network can use
another docker container's name as the hostname and communicate!
``` {r }
nw <- docker$network$create("mynetwork")
server <- docker$container$run("nginx", network = nw, name = "server",
                               detach = TRUE, rm = TRUE)
server$status()
```

Now we can attach other networks to this container and communicate
with the server:
``` {r }
docker$container$run("alpine:latest", c("ping", "server", "-c", "3"),
                     network = nw, stream = stdout(), rm = TRUE)
```

Omitting the `network` argument, the second container can't find
the server:
``` {r error = TRUE}
docker$container$run("alpine:latest", c("ping", "server", "-c", "3"),
                     stream = stdout(), rm = TRUE)
```

The server container exposes a webserver on port 80.  For
containers on the network we can access this port:
``` {r }
res <- docker$container$run("richfitz/curl", c("curl", "-s", "http://server"),
                            network = nw, rm = TRUE)
head(res$logs, 10)

server$stop()
nw$remove()
```

## Other docker functions

There are a few functions at the top level of the `docker_client`
object:

`$ping()` tests the connection to the server and reports the API
version for the server - this is a (for docker) very fast function
to use to test that things seem to be working.
``` {r }
docker$ping()
```

`$api_version()` reports the API version that the *client* is using
(this can be varied from `r stevedore:::DOCKER_API_VERSION_MIN` to
`r stevedore:::DOCKER_API_VERSION_MAX`)
``` {r }
docker$api_version()
```

`$version()` reports detailed version information from the server:
``` {r }
docker$version()
```

`$info()` reports a bunch of other information about the state of
the server (Docker describes this as "get system information" in
its documentation) - this is equivalent to running `docker info`
``` {r }
docker$info()
```

Finally, `$df()` will return information about resource and data
usage by docker - all containers, networks, volumes, etc.
